---
title: "Species Climate Exposure Dashboard"
format:
  dashboard:
    nav-buttons:
      - icon: github
        href: https://github.com/cmerow/2025_Exposure
        aria-label: GitHub
---

```{=html}
<style>
/* Prevent card focus state from resizing the map card */
.card:focus-within,
.bslib-card:focus-within {
  outline: none !important;
  box-shadow: none !important;
}
/* Prevent any flex child from growing beyond its allocated height.
   This stops the species table card from pushing the map row up.
   Every element in the flex chain must have min-height: 0. */
.bslib-gap-spacing,
.bslib-gap-spacing > *,
.html-fill-container,
.html-fill-item,
.bslib-card,
.bslib-card .card-body {
  min-height: 0 !important;
  min-width: 0 !important;
}
/* Cards in the bottom row scroll internally rather than expand */
.bslib-card .card-body {
  overflow: auto !important;
  overflow-x: hidden !important;
}
/* Force the leaflet widget and its wrappers to fill the card completely */
.bslib-card .card-body .html-widget,
.bslib-card .card-body .html-widget-output,
.bslib-card .card-body .leaflet,
.bslib-card .card-body .leaflet-container {
  width: 100% !important;
  height: 100% !important;
}
</style>
```

```{r setup, include=FALSE}
# =============================================================================
# Setup chunk – loaded once at render time.
#
# Reads the pre-computed cache files produced by _R/01_prepare-data.R.
# No heavy computation happens here; all summarisation was done at build time.
# =============================================================================

library(dplyr)
library(ggplot2)
library(plotly)
library(leaflet)
library(jsonlite)
library(leaflet.extras)
library(DT)
library(terra)
library(raster)   # addRasterImage expects a {raster} object

# Source plotting helpers
source("_R/timeseries-plotting.R")

# ---------------------------------------------------------------------------
# Load pre-computed cache files
# These are written by _R/01_prepare-data.R during the build step.
# ---------------------------------------------------------------------------

# Main exposure data: full table + year index + trend tables
dashboard_data <- readRDS("data/exposure-dashboard.rds")

exposure_df   <- dashboard_data$exposure_df
year_index    <- dashboard_data$year_index
cell_trend    <- dashboard_data$cell_trend
sp_trend      <- dashboard_data$sp_trend
years_avail   <- dashboard_data$years_avail
vars_avail    <- dashboard_data$vars_avail
species_avail <- dashboard_data$species_avail

# Species metadata (name, order, family, IUCN status)
species_metadata <- readRDS("data/species-metadata.rds")

# Per-species timeseries cache (keyed by species name)
ts_cache <- readRDS("data/species-timeseries-cache.rds")

# Land template raster (defines the spatial grid; cells match exposure_df$cell)
# Loaded from data/ where the build step should have copied it from the release
tpl <- terra::rast("data/landTemplate.tif")

# ---------------------------------------------------------------------------
# Shared constants
# ---------------------------------------------------------------------------

# Human-readable labels for climate variable codes
VAR_LABELS <- c(
  temp__12_up       = "Temperature: annual upper extreme",
  temp__12_lo       = "Temperature: annual lower extreme",
  temp__3__max_up   = "Temperature: seasonal max upper extreme",
  temp__3__min_lo   = "Temperature: seasonal min lower extreme",
  precip__12_up     = "Precipitation: annual upper extreme",
  precip__12_lo     = "Precipitation: annual lower extreme",
  precip__3__max_up = "Precipitation: seasonal max upper extreme",
  precip__3__min_lo = "Precipitation: seasonal min lower extreme"
)

# Colours for each climate variable (for map overlays and legends)
VAR_COLORS <- c(
  temp__12_up       = "#d73027",
  temp__12_lo       = "#4575b4",
  temp__3__max_up   = "#f46d43",
  temp__3__min_lo   = "#313695",
  precip__12_up     = "#1a9850",
  precip__12_lo     = "#9400D3",
  precip__3__max_up = "#33a02c",
  precip__3__min_lo = "#6a0dad"
)

# Helper: translate a variable code to its display label
var_label <- function(v) {
  lab <- VAR_LABELS[as.character(v)]
  dplyr::if_else(is.na(lab), as.character(v), lab)
}

# Helper: translate a variable code to its map colour
var_colour <- function(v) {
  col <- VAR_COLORS[as.character(v)]
  dplyr::if_else(is.na(col), "#e41a1c", col)
}

# Most recent year in the data (data always lags by ~1 year)
RECENT_YEAR   <- max(years_avail)           # e.g. 2025
RECENT_CUTOFF <- 2023L                      # first year of "Recent observed" period
hotspot_years <- RECENT_CUTOFF:RECENT_YEAR  # years shown in Hotspot Explorer

# Default selections shown on first load
default_year <- RECENT_YEAR
default_species <- species_avail[1]

# Named vector for the variable checkbox: label → code
var_choices <- setNames(vars_avail, vapply(vars_avail, var_label, character(1)))
```

# Species Explorer {.tabset}

## {.sidebar}

**Species Explorer**

Select a species and a subset of climate variables to see how much of its
range has been exposed to extreme climate events through time.

```{r species-sidebar, echo=FALSE}
# Sidebar controls are standard Quarto/HTML inputs connected to
# ojs (Observable JS) reactive blocks below via `ojs_define()`.

# Pass R vectors to the OJS layer so controls can be populated dynamically
ojs_define(
  species_list       = species_avail,
  vars_list          = unname(var_choices),
  var_names          = names(var_choices),
  years_list         = years_avail,
  hotspot_years_list = hotspot_years,
  default_year       = default_year
)
```

```{ojs}
//| echo: false
// Species selector
viewof selected_species = Inputs.select(
  species_list,
  { value: species_list[0], label: "Species" }
)
```

```{ojs}
//| echo: false
// Climate variable checkboxes
viewof selected_vars = Inputs.checkbox(
  var_names,
  {
    value: var_names,   // all selected by default
    label: "Climate variables"
  }
)
```

---

*Data: historical observations up to 2025*

## Species

### Row {height="28%"}

#### Species info {width="45%"}

```{r species-meta-table, echo=FALSE}
species_display <- species_metadata |>
  dplyr::rename(
    Species       = spName,
    Order         = dplyr::any_of("orderName"),
    Family        = dplyr::any_of("familyName"),
    `IUCN status` = dplyr::any_of("redlistCategory")
  )
ojs_define(species_meta_data = species_display)
```

```{ojs}
//| echo: false
filtered_meta = transpose(species_meta_data).filter(row => row.Species === selected_species)
Inputs.table(filtered_meta, { width: "100%" })
```

#### Range map {width="55%"}

```{ojs}
//| echo: false
species_range_map = {
  const L = await require("https://unpkg.com/leaflet@1.9.4/dist/leaflet.js");

  const container = DOM.element("div");
  // Fixed pixel height required: L.map() measures the container at init time,
  // before OJS inserts it into the DOM, so height:100% would resolve to 0.
  container.style.cssText = "width:100%;height:280px;";

  const sp = encodeURIComponent(selected_species);
  let mapData;
  try {
    const resp = await fetch(`data/species_maps/${sp}_map.json`);
    if (!resp.ok) throw new Error(`HTTP ${resp.status}`);
    mapData = await resp.json();
  } catch(e) {
    const msg = DOM.element("p");
    msg.textContent = "Range map not yet built \u2014 run the build step.";
    msg.style.cssText = "color:#999;font-size:12px;text-align:center;margin-top:2em;";
    return msg;
  }

  const map = L.map(container, { preferCanvas: true });
  L.tileLayer("https://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}{r}.png", {
    attribution: "\u00a9 OpenStreetMap, \u00a9 CARTO",
    maxZoom: 19
  }).addTo(map);
  // Re-measure after OJS inserts the container into the DOM
  setTimeout(() => map.invalidateSize(), 100);

  const renderer = L.canvas({ padding: 0.5 });
  const bounds = [];

  // Colour: 1 variable = blue, 8 variables = red
  const varColor = (v) => {
    const t = Math.min((v - 1) / 7, 1);
    const r = Math.round(t * 215 + (1 - t) * 49);
    const g = Math.round((1 - t) * 115);
    const b = Math.round((1 - t) * 178 + t * 39);
    return `rgb(${r},${g},${b})`;
  };

  for (const [c, v] of mapData) {
    const col = (c - 1) % rast_ncols + 1;
    const row = Math.floor((c - 1) / rast_ncols) + 1;
    const lon = rast_xmin + (col - 0.5) * rast_xres;
    const lat = rast_ymax - (row - 0.5) * rast_yres;
    L.circleMarker([lat, lon], {
      radius: 4,
      fillColor: varColor(v),
      color: "none",
      fillOpacity: 0.75,
      renderer
    }).addTo(map);
    bounds.push([lat, lon]);
  }

  if (bounds.length)
    map.fitBounds(L.latLngBounds(bounds), { padding: [10, 10] });

  invalidation.then(() => map.remove());
  return container;
}
```

### Row {height="35%"}

```{r species-figures-data, echo=FALSE}
# Species figures are served as static PNG files under data/species_figs/.
# OJS fetches them lazily by constructing the URL from the selected species name.
ojs_define(has_species_figs = file.exists("data/species_figs"))
```

```{ojs}
//| echo: false
{
  const sp  = encodeURIComponent(selected_species);
  const base = `data/species_figs/${sp}`;
  const wrap = DOM.element("div");
  wrap.style.cssText = "display:flex;width:100%;height:100%;gap:8px;overflow:hidden;";

  const makeImg = (src, title) => {
    const box = DOM.element("div");
    box.style.cssText = "flex:1;display:flex;flex-direction:column;align-items:center;min-width:0;";
    const img = DOM.element("img");
    img.src = src;
    img.style.cssText = "max-width:100%;max-height:100%;object-fit:contain;";
    img.onerror = () => {
      const msg = DOM.element("p");
      msg.textContent = title + " not yet built \u2014 run the build step.";
      msg.style.cssText = "color:#999;font-size:12px;text-align:center;margin-top:2em;";
      box.replaceChildren(msg);
    };
    box.appendChild(img);
    return box;
  };

  wrap.appendChild(makeImg(`${base}_kde.png`,   "KDE figure"));
  wrap.appendChild(makeImg(`${base}_polar.png`, "Polar figure"));
  return wrap;
}
```

### Row {height="20%"}

```{r exposure-table-all, echo=FALSE}
all_summaries <- purrr::imap_dfr(ts_cache, function(sp_data, sp_name) {
  sp_data$summary |>
    dplyr::filter(n_cells_exposed > 0) |>
    dplyr::mutate(
      species           = sp_name,
      mean_prop_exposed = round(mean_prop_exposed, 4)
    )
})
ojs_define(all_summaries_data = all_summaries)
```

```{ojs}
//| echo: false
filtered_summary = transpose(all_summaries_data).filter(d => d.species === selected_species)

Inputs.table(
  filtered_summary.map(({ year, period, n_cells_exposed, mean_prop_exposed }) => ({
    Year: String(year),
    Period: period,
    "Cells exposed": String(n_cells_exposed),
    "Mean % range exposed": (mean_prop_exposed * 100).toFixed(2) + "%"
  })),
  { width: "100%", sort: "Year", reverse: true }
)
```

# Hotspot Explorer {scrolling="false"}

## {.sidebar}

**Hotspot Explorer**

Click any cell on the map to see which species are exposed there.

```{ojs}
//| echo: false
viewof hotspot_year = Inputs.select(
  hotspot_years_list.map(String),
  { value: String(default_year), label: "Year" }
)
```

```{ojs}
//| echo: false
viewof hotspot_vars = Inputs.checkbox(
  var_names,
  { value: var_names, label: "Climate variables" }
)
```

```{ojs}
//| echo: false
viewof hotspot_threatened_only = Inputs.toggle({
  label: "Only threatened species (CR / EN / VU)",
  value: false
})
```

## {.fill}

### Map {.fill}

```{r hotspot-map, echo=FALSE}
#| fill: true
# Build an interactive Leaflet map showing the hotspot raster.
# All years are pre-rendered as stacked image layers; JS toggles visibility.

# Build rasters for every available year using a shared colour domain
# so the colour scale is consistent across years.
all_hotspot_counts <- exposure_df |>
  dplyr::filter(!is.na(cell), cell >= 1, cell <= terra::ncell(tpl)) |>
  dplyr::distinct(spName, cell, year) |>
  dplyr::count(cell, year, name = "n_species")

global_vals <- all_hotspot_counts$n_species
global_pal  <- leaflet::colorNumeric(
  palette   = colorRampPalette(c("steelblue4", "steelblue1", "gold", "red1", "red4"))(100),
  domain    = log10(c(1, max(global_vals))),
  na.color  = "transparent"
)
# Wrapper: maps raw counts → log10 → colour
pal_fn <- function(x) global_pal(log10(pmax(x, 1)))
# Legend palette and values on log10 scale for readability
legend_breaks <- c(1, 2, 5, 10, 20, 50, 100, max(global_vals))
legend_breaks <- legend_breaks[legend_breaks <= max(global_vals)]

# Start map
lmap <- leaflet::leaflet(height = "100%", options = leaflet::leafletOptions(preferCanvas = TRUE)) |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.DarkMatter) |>
  leaflet::setView(lng = 0, lat = 15, zoom = 2)

# Add one raster overlay per hotspot year (2023–present); only default_year starts visible
for (yr in hotspot_years) {
  yr_counts <- all_hotspot_counts |> dplyr::filter(year == yr)
  rv <- rep(NA_real_, terra::ncell(tpl))
  rv[yr_counts$cell] <- yr_counts$n_species
  yr_raster <- raster::raster(terra::setValues(tpl, rv))
  lmap <- lmap |> leaflet::addRasterImage(
    yr_raster,
    colors  = pal_fn,
    opacity = if (yr == default_year) 1 else 0,
    layerId = paste0("hotspot_", yr)
  )
}

# Add legend and click handler
default_year_df <- exposure_df[year_index[[as.character(default_year)]], ]
lmap <- lmap |>
  leaflet::addLegend(
    position = "bottomright",
    colors   = global_pal(log10(pmax(legend_breaks, 1))),
    labels   = as.character(legend_breaks),
    title    = "# Species exposed",
    opacity  = 1
  ) |>
  # Wire map clicks to a custom JS event so OJS can react.
  # Use paste0 so hotspot_years is embedded as a JS array literal at render time.
  htmlwidgets::onRender(paste0("
    function(el, x) {
      var map = this;
      map.invalidateSize(false);
      setTimeout(function() { map.invalidateSize(false); }, 300);
      // Listen for year changes from OJS and toggle raster layer opacity.
      // addRasterImage registers layers via layerManager (not layer.options),
      // so we must use map.layerManager.getLayer(id, category) to access them.
      var hotspotYears = [", paste(hotspot_years, collapse = ","), "];
      window.addEventListener('hotspot_year_change', function(e) {
        var yr = String(e.detail.year);
        hotspotYears.forEach(function(y) {
          var layer = map.layerManager.getLayer('hotspot_' + y, 'image');
          if (layer) layer.setOpacity(String(y) === yr ? 1 : 0);
        });
      });
      map.on('click', function(e) {
        // Update or add a red circle marker at the clicked location
        if (window._clickMarker) { window._clickMarker.remove(); }
        window._clickMarker = L.circleMarker([e.latlng.lat, e.latlng.lng], {
          radius: 8, color: '#ff2222', weight: 2,
          fillColor: '#ff2222', fillOpacity: 0.5
        }).addTo(map);
        window._mapClick = { lat: e.latlng.lat, lng: e.latlng.lng, t: Date.now() };
        window.dispatchEvent(new CustomEvent('leaflet_mapclick', { detail: window._mapClick }));
      });
    }
  "))
bslib::as_fill_item(lmap)
```

```{r raster-geometry, include=FALSE}
# Pass raster geometry to OJS so the click handler can convert lat/lng → cell index.
# Terra cell numbering: row-major, 1-indexed, row 1 = northernmost.
rast_ext  <- as.vector(terra::ext(tpl))   # xmin, xmax, ymin, ymax
rast_res  <- terra::res(tpl)              # xres, yres (both positive)
ojs_define(
  rast_xmin  = rast_ext[1],
  rast_ymax  = rast_ext[4],
  rast_xres  = rast_res[1],
  rast_yres  = rast_res[2],
  rast_ncols = terra::ncol(tpl)
)
```

```{r cell-lookup-data, include=FALSE}
# Write compact per-year JSON files + small lookup files.
# Format: [{c: cell, s: spName, v: varIdx, p: propExposed, r: redlistIdx}]
# Lookup files (written once): var_codes.json, redlist_codes.json, species_attrs.json
# This reduces file size ~75% vs the old verbose format.
cell_species_dir <- "data/cell_species"
dir.create(cell_species_dir, recursive = TRUE, showWarnings = FALSE)

# --- build lookup tables ---
all_var_labels <- sort(unique(vapply(as.character(unique(exposure_df$var)), var_label, character(1))))
var_idx        <- setNames(seq_along(all_var_labels) - 1L, all_var_labels)
jsonlite::write_json(as.list(all_var_labels),
                     file.path(cell_species_dir, "var_codes.json"), auto_unbox = TRUE)

all_redlist  <- sort(unique(as.character(exposure_df$redlistCategory)))
redlist_idx  <- setNames(seq_along(all_redlist) - 1L, all_redlist)
jsonlite::write_json(as.list(all_redlist),
                     file.path(cell_species_dir, "redlist_codes.json"), auto_unbox = TRUE)

sp_attrs <- dplyr::distinct(exposure_df, spName, orderName, familyName) |>
  dplyr::filter(!is.na(spName))
sp_attrs_list <- setNames(
  lapply(seq_len(nrow(sp_attrs)), function(i)
    list(o = sp_attrs$orderName[i], f = sp_attrs$familyName[i])),
  sp_attrs$spName)
jsonlite::write_json(sp_attrs_list,
                     file.path(cell_species_dir, "species_attrs.json"), auto_unbox = TRUE)

# --- per-year compact files (Hotspot Explorer: recent years only) ---
for (yr in RECENT_CUTOFF:RECENT_YEAR) {
  yr_path <- file.path(cell_species_dir, paste0("species_", yr, ".json"))
  if (!file.exists(yr_path)) {
    yr_data <- exposure_df[year_index[[as.character(yr)]], ] |>
      dplyr::filter(!is.na(cell)) |>
      dplyr::mutate(
        vl = vapply(as.character(var), var_label, character(1)),
        v  = var_idx[vl],
        p  = round(as.numeric(propExposed), 4),
        r  = redlist_idx[as.character(redlistCategory)]
      ) |>
      dplyr::arrange(cell, v) |>
      dplyr::select(c = cell, s = spName, v, p, r)
    jsonlite::write_json(yr_data, yr_path, auto_unbox = TRUE)
  }
}
```

```{ojs}
//| output: false
// Listen for leaflet_mapclick events dispatched by the onRender handler.
// Converts lat/lng to a terra cell index (row-major, 1-indexed).
clicked_cell = Generators.observe(notify => {
  function handler(e) {
    const { lat, lng } = e.detail;
    if (lng < rast_xmin || lng > (rast_xmin + rast_ncols * rast_xres) ||
        lat > rast_ymax || lat < (rast_ymax - rast_ncols * rast_yres)) {
      return; // outside raster extent — ignore
    }
    const col  = Math.floor((lng - rast_xmin) / rast_xres) + 1;
    const row  = Math.floor((rast_ymax - lat)  / rast_yres) + 1;
    const cell = (row - 1) * rast_ncols + col;
    notify(String(cell));
  }
  window.addEventListener('leaflet_mapclick', handler);
  notify(""); // initial: nothing selected
  return () => window.removeEventListener('leaflet_mapclick', handler);
})
```

```{ojs}
//| output: false
// Dispatch year change events to the map when hotspot_year changes
{
  window.dispatchEvent(new CustomEvent('hotspot_year_change', { detail: { year: hotspot_year } }));
}
```

### {height="320px"}

#### Species in selected cell {height="100%"}

```{ojs}
//| output: false
// Compute formatted lat/lon from clicked cell index
click_latlon = {
  if (clicked_cell === "") return "";
  const c = Number(clicked_cell);
  const col = (c - 1) % rast_ncols + 1;
  const row = Math.floor((c - 1) / rast_ncols) + 1;
  const lonRaw = rast_xmin + (col - 0.5) * rast_xres;
  const latRaw = rast_ymax - (row - 0.5) * rast_yres;
  const lon = Math.abs(lonRaw).toFixed(2) + "\u00b0" + (lonRaw >= 0 ? "E" : "W");
  const lat = Math.abs(latRaw).toFixed(2) + "\u00b0" + (latRaw >= 0 ? "N" : "S");
  return `${lat}, ${lon}`;
}
```

```{ojs}
//| output: false
// Load the three small lookup files once (not re-fetched when year changes).
cell_species_lookups = Promise.all([
  fetch("data/cell_species/var_codes.json").then(r => r.json()),
  fetch("data/cell_species/redlist_codes.json").then(r => r.json()),
  fetch("data/cell_species/species_attrs.json").then(r => r.json()),
])
```

```{ojs}
//| output: false
// Fetch the species exposure JSON for the selected year lazily.
// Only the current year's compact file is loaded; others are fetched on demand.
cell_species = {
  const raw = await fetch(`data/cell_species/species_${hotspot_year}.json`).then(r => r.json());
  const [varCodes, redlistCodes, spAttrs] = await cell_species_lookups;
  return raw.map(d => ({
    cell:            d.c,
    spName:          d.s,
    var:             varCodes[d.v],
    propExposed:     d.p,
    redlistCategory: redlistCodes[d.r] ?? "—",
    orderName:       spAttrs[d.s]?.o   ?? "—",
    familyName:      spAttrs[d.s]?.f   ?? "—"
  }));
}
```

```{ojs}
//| output: false
cell_filtered = clicked_cell === ""
  ? []
  : cell_species
      .filter(d => d.cell === Number(clicked_cell))
      .filter(d => hotspot_vars.includes(d.var))
      .filter(d => !hotspot_threatened_only || ["CR", "EN", "VU"].includes(d.redlistCategory))
```

```{ojs}
//| echo: false
clicked_cell === ""
  ? html`<p style="color:#888;font-style:italic">Click any location on the map to see which species are exposed there.</p>`
  : html`<div>
    <h6 style="margin:0 0 6px">Species exposed at this location in ${hotspot_year} (${click_latlon})</h6>
    ${Inputs.table(
      cell_filtered.map(({ spName, var: v, propExposed, orderName, familyName, redlistCategory }) => ({
        Species:           spName,
        Variable:          v,
        "% range exposed": propExposed ? (propExposed * 100).toFixed(2) + "%" : "—",
        Order:             orderName   ?? "—",
        Family:            familyName  ?? "—",
        "IUCN status":     redlistCategory ?? "—"
      })),
      { width: "100%", sort: "% range exposed", reverse: true }
    )}
    </div>`
```

#### Cell trend over time {height="100%"}

```{r cell-trend-data, include=FALSE}
# Build per-variable cell trend (species count per cell × year × variable).
# Computed directly from exposure_df so we don't need to re-run prepare-data.
cell_trend_by_var <- exposure_df |>
  dplyr::filter(!is.na(cell), cell >= 1, cell <= terra::ncell(tpl)) |>
  dplyr::mutate(var = vapply(as.character(var), var_label, character(1))) |>
  dplyr::distinct(spName, cell, year, var) |>
  dplyr::count(cell, year, var, name = "n_species_exposed")
ojs_define(cell_trend_data = cell_trend_by_var)
```

```{ojs}
//| output: false
cell_trend_filtered = {
  if (clicked_cell === "") return [];
  const raw = transpose(cell_trend_data)
    .filter(d => d.cell === Number(clicked_cell))
    .filter(d => hotspot_vars.includes(d.var));
  // Build lookup: "year|var" -> n_species_exposed
  const lookup = new Map(raw.map(d => [`${d.year}|${d.var}`, d.n_species_exposed]));
  // Cross-join recent years (>2010) × selected vars, filling missing with 0
  const allYears = years_list.map(Number).filter(y => y > 2010);
  const rows = [];
  for (const year of allYears) {
    for (const v of hotspot_vars) {
      rows.push({ year, var: v, n_species_exposed: lookup.get(`${year}|${v}`) ?? 0 });
    }
  }
  return rows;
}
```

```{ojs}
//| echo: false
{
  if (clicked_cell === "" || cell_trend_filtered.every(d => d.n_species_exposed === 0))
    return html`<p style="color:#888;font-style:italic">Click any location on the map to see which species are exposed there.</p>`;
  const c = Number(clicked_cell);
  const col = (c - 1) % rast_ncols + 1;
  const row = Math.floor((c - 1) / rast_ncols) + 1;
  const lonRaw = rast_xmin + (col - 0.5) * rast_xres;
  const latRaw = rast_ymax - (row - 0.5) * rast_yres;
  const lon = Math.abs(lonRaw).toFixed(2) + "\u00b0" + (lonRaw >= 0 ? "E" : "W");
  const lat = Math.abs(latRaw).toFixed(2) + "\u00b0" + (latRaw >= 0 ? "N" : "S");
  return Plot.plot({
        title: `Number of species exposed per variable (${click_latlon})`,
        width,
        height: 260,
        marginBottom: 50,
        style: { fontSize: "14px" },
        fx: { label: "Year", tickFormat: d => d },
        x:  { axis: null },
        y:  { label: "Species exposed", grid: true },
        color: { legend: true },
        marks: [
          Plot.barY(cell_trend_filtered, {
            fx: d => String(d.year),
            x: "var",
            y: "n_species_exposed",
            fill: "var",
            tip: true
          }),
          Plot.ruleY([0])
        ]
      });
}
```
